---
title: "Mortgage Interest Rate and Consumption"
output: 
  html_document:
    css: bodycss.css
    fig_width: 8
    font-family: Helvetica,Arial,sans-serif;
    number_section: yes
    toc: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

<style type="text/css">
body{ 
      font-size: 15px;
      font-family: Helvetica,Arial,sans-serif;
      line-height: 200%;
  }
  
.author {
 font-size: 15px;
 color: Black;
 font-style: normal;
 text-align: center;
}


.date { 
 font-size: 15px;
 color: Black;
 font-style: normal;
 text-align: center;
}

.title{
  text-align: center;
  font-size: 15px;
 color: Black;
 
}

.toc-ignore{
  text-align: center;
  font-size: 15px;
 color: Black;
}

.fluid-row{
  text-align: center;
  font-size: 15px;
 color: Black;
}

</style>

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE,echo=FALSE)
```

```{r echo=FALSE}
rm(list=ls())


library(data.table)
library(lubridate)
library(lfe)
library(plyr)
library(stargazer)
library(FinCal)
library(ggplot2)
output.type = "text"

completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}

merged_data <- readRDS(file="E:/CE Survey/Data/Processed/merged_data_Nov302017_CQ.rds") # file created by 11012017_kevin_comments.R





merged_data['housing_exp_calc'] <- merged_data$major_appliances+merged_data$small_appliances + merged_data$maintainance_repairs+merged_data$house_furnishings_and_equip+merged_data$mtg_principal_outlay

merged_data['non_durable'] <- merged_data$food_exp+ merged_data$entertainment_exp + merged_data$utility_exp + merged_data$transport_outlays - merged_data$vehicle_purchase_outlay + merged_data$apperel_and_services_exp+merged_data$miscellaneous_exp 

merged_data['durable'] <- merged_data$major_appliances + merged_data$small_appliances + merged_data$vehicle_purchase_outlay +merged_data$tv_radio_etc

merged_data['services'] <- merged_data$education_exp + merged_data$domestic_services + merged_data$health_exp


merged_data$family_income_after_tax <- merged_data$family_income_after_tax+1

income_medians <- ddply(merged_data,.(state,interview_year),summarise,q1_income = quantile(family_income_after_tax,0.25,na.rm=TRUE),median_income = median(family_income_after_tax,na.rm=TRUE),q3_income = quantile(family_income_after_tax,0.75,na.rm=TRUE))

merged_data <- merge(merged_data,income_medians,by=c("state","interview_year"),all.x = TRUE)

merged_data$mtg30us <- merged_data$mtg30us*100

merged_data <- completeFun(merged_data,c("family_income_after_tax","family_size","age_of_ref",
                                  "interview_state","mpay_mtg","mtg30us","CUID","org_state","state"))
t1 <- as.data.frame(table(merged_data$org_state))
t2 <- as.data.frame(table(merged_data$interview_state))

merged_data['org_state_int_state'] <- paste(merged_data$org_state,merged_data$interview_state)
t3 <- as.data.frame(table(merged_data$org_state_int_state))

merged_data <- merged_data[!merged_data$org_state %in% t1[t1$Freq==1,]$Var1,]
merged_data <- merged_data[!merged_data$interview_state %in% t2[t2$Freq==1,]$Var1,]
merged_data <- merged_data[!merged_data$org_state_int_state %in% t3[t3$Freq==1,]$Var1,]

merged_data['principal_reduction'] <- merged_data$mtg_pmt_calc_mor*3 - (as.numeric(merged_data$QADINT1X) + as.numeric(merged_data$QADINT2X) + as.numeric(merged_data$QADINT3X) + as.numeric(merged_data$QPRINM1X) + as.numeric(merged_data$QPRINM2X) + as.numeric(merged_data$QPRINM3X))
merged_data$principal_reduction <- ifelse(merged_data$principal_reduction<0,0,merged_data$principal_reduction)


merged_data$non_durable <- ifelse(merged_data$int_month==2 | merged_data$int_month==5 | merged_data$int_month==8 | merged_data$int_month==11,merged_data$non_durable*3, merged_data$non_durable*3/2)
merged_data$durable <- ifelse(merged_data$int_month==2 | merged_data$int_month==5 | merged_data$int_month==8 | merged_data$int_month==11,merged_data$durable*3, merged_data$durable*3/2)
merged_data$services <- ifelse(merged_data$int_month==2 | merged_data$int_month==5 | merged_data$int_month==8 | merged_data$int_month==11,merged_data$services*3, merged_data$services*3/2)
merged_data$total_exp <- ifelse(merged_data$int_month==2 | merged_data$int_month==5 | merged_data$int_month==8 | merged_data$int_month==11,merged_data$total_exp*3, merged_data$total_exp*3/2)
# merged_data$mtg_pmt_reported_mor <- ifelse(merged_data$int_month==2 | merged_data$int_month==5 | merged_data$int_month==8 | merged_data$int_month==11,merged_data$mtg_pmt_reported_mor*3, merged_data$mtg_pmt_reported_mor*3/2)
merged_data['total_mtg_pmt'] <- merged_data$mtg_pmt_reported_mor*3

merged_data$interest_rate_cur <- merged_data$interest_rate_cur*100
```

# Descriptives
```{r descs}

stargazer(merged_data[, c("interest_rate_cur","mtg30us","mtg_amount","non_durable","durable","services","total_mtg_pmt","total_exp","mtg_principal_outlay","org_year","interview_year","family_income_after_tax","family_size","age_of_ref")], type = output.type, summary.stat = c("mean", "sd", "p25", "median", "p75", "n"),notes = "",digits = 4)
```
```{r include=FALSE}
stargazer(merged_data[, c("total_exp","mtg_amount","food_exp","entertainment_exp","utility_exp","transport_outlays","apperel_and_services_exp","miscellaneous_exp","major_appliances","small_appliances","tv_radio_etc","vehicle_purchase_outlay","education_exp","domestic_services","health_exp","mtg_principal_outlay","maintainance_repairs","house_furnishings_and_equip","pets_toys","life_and_personal_insurance")], type = output.type, summary.stat = c("mean", "sd", "p25", "median", "p75", "n"),notes = "",digits = 4)
```

# Balancing Tests
```{r balancing}
formula_text <- "~0|org_state+interview_state|(interest_rate_org~mtg30us)|interview_year+CUID"
all_interest_data <- readRDS(file="E:/CE Survey/Data/Processed/merged_data_all_interestrate_Nov142017.rds")
all_interest_data <- all_interest_data[all_interest_data$org_int_not_eq_cur==1 | all_interest_data$NEWID.x %in% unique(merged_data$NEWID.x),]

variables <- c("log(family_income_after_tax)","age_of_ref","family_size","log(mtg_amount)")
i = 1
felm_regs <- list()
for(variable in variables) {
  # print(i)
  felm_regs[[i]] <- felm(as.formula(paste(variable,formula_text,sep="")),data=merged_data[is.finite(log(merged_data$family_income_after_tax)),])
  i=i+1
}
felm_regs[[i]] <- felm(as.formula(paste("org_int_not_eq_cur",formula_text,sep="")),data=all_interest_data[is.finite(log(all_interest_data$family_income_after_tax)),])

stargazer(felm_regs,type = output.type,no.space = TRUE,column.labels = gsub("_"," ",c(variables,"org_int_not_eq_cur")),dep.var.labels.include = FALSE)
```

# Kernal Regression
```{r kernal_reg}

library(np)
temp <- merged_data

temp <- completeFun(temp,c("mtg30us","interest_rate_org","org_state","interview_state"))

adj_mpay_mtg <- felm(interest_rate_org~0|factor(org_state)+factor(interview_state),data=temp)
adj_mtg30us <- felm(mtg30us~0|factor(org_state)+factor(interview_state),data=temp)
temp['resid_mpay_mtg'] <- adj_mpay_mtg$residuals
temp['resid_mtg30us'] <- adj_mtg30us$residuals

bw <- npregbw(formula=as.vector(temp$resid_mpay_mtg)~as.vector(temp$resid_mtg30us),bws=(max(temp$resid_mtg30us)-min(temp$resid_mtg30us))/10,bwtype="fixed",bandwidth.compute=F)

model <- npreg(bws=bw,gradients = TRUE)
plot.out <- plot(model, plot.errors.method="asymptotic",plot.errors.style="band",plot.behavior="data")
y.eval <- fitted(plot.out$r1)
x.eval <- plot.out$r1$eval[,1]
y.se <- se(plot.out$r1)
y.lower.ci <- y.eval+1.96*y.se[,1]
y.upper.ci <- y.eval+1.96*y.se[,2]
df <- as.data.frame(cbind(x.eval,y.eval,y.lower.ci,y.upper.ci))

df <- melt(df,id="x.eval")

g2 <- ggplot(df,aes(x=x.eval, y=value,colour=variable)) + geom_line(aes(linetype=variable), size=1) +scale_linetype_manual(values = c(1,2,2))+scale_colour_manual(values=c("black","gray40","gray40"))+ theme_bw()+ylab("FE adjusted Interest Rate") + xlab("FE adjusted IV")+ labs(title = "")+ theme(legend.position="none")
g2



```




# Regressions for Paper
Columns
1 - First Stage <br/>
2 - OLS - non_durable<br/>
3 - IV - non_durable<br/>
4 - OLS - durable<br/>
5 - IV - durable<br/>
6 - OLS - services<br/>
7 - IV - services<br/>
8 - OLS - principal_reduction<br/>
9 - IV - principal_reduction<br/>
8 - OLS - Mortgage Payment<br/>
9 - IV - Mortgage Payment<br/>
10 - OLS - total_exp<br/>
11 - IV - total_exp<br/>
```{r regs_for_paper}

reg_data <- merged_data[is.finite(log(merged_data$family_income_after_tax)) & !is.na(merged_data$state),]


endo_var = "interest_rate_cur"
instrument = "mtg30us"
cluster_var = "CUID+interview_year"

controls <- "log(family_income_after_tax)+family_size+I(age_of_ref^2)|org_state+interview_state"

dep_vars <- c("non_durable","durable","services","principal_reduction","total_mtg_pmt","total_exp")
regs <- list()


regs[[1]] <- felm(as.formula(paste("I(",endo_var,")~",instrument,"+",controls,"|0|",cluster_var,sep="")),data = reg_data)
i=2
for(dep_var in dep_vars) {
  
  reg_data <- merged_data[is.finite(log(merged_data$family_income_after_tax)) & !is.na(merged_data$state),]
  eval(parse(text=paste("reg_data$",dep_var," <- abs(reg_data$",dep_var,"+1)",sep="")))
  eval(parse(text=paste("reg_data$",dep_var," <- ifelse(is.na(reg_data$",dep_var,"),1,reg_data$",dep_var,")",sep="")))
  
  
  log_dep_var = paste("log(",dep_var,")",sep="")
  regs[[i]] <- felm(as.formula(paste(log_dep_var,"~",endo_var,"+",controls,"|0|",cluster_var,sep="")),data = reg_data)
  i=i+1
  regs[[i]] <- felm(as.formula(paste(log_dep_var,"~",controls,"|(",endo_var,"~",instrument,")|",cluster_var,sep="")),data = reg_data)
  i=i+1
  
}

stargazer(regs,type="text",dep.var.labels.include = FALSE,add.lines = list(c("Cond. F","","",round(condfstat(regs[[3]])[[1]],2),"",round(condfstat(regs[[5]])[[1]],2),""
                               ,round(condfstat(regs[[7]])[[1]],2),"",round(condfstat(regs[[9]])[[1]],2),"",round(condfstat(regs[[11]])[[1]],2))))

```
