---
title: "Mortgage Payment Size and Consumption"
output: 
  html_document: 
    fig_height: 6
    fig_width: 12
    toc: yes
    number_section: yes
---

<style type="text/css">
body{ 
      font-size: 15px;
      font-family: Helvetica,Arial,sans-serif;
      line-height: 200%;
  }
  
.author {
 font-size: 15px;
 color: Black;
 font-style: normal;
 text-align: center;
}


.date { 
 font-size: 15px;
 color: Black;
 font-style: normal;
 text-align: center;
}

.title{
  text-align: center;
  font-size: 15px;
 color: Black;
 
}

.toc-ignore{
  text-align: center;
  font-size: 15px;
 color: Black;
}

.fluid-row{
  text-align: center;
  font-size: 15px;
 color: Black;
}

</style>

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE,echo=FALSE)
```

```{r echo=FALSE}
rm(list=ls())


library(data.table)
library(lubridate)
library(lfe)
library(plyr)
library(stargazer)
library(FinCal)
library(ggplot2)
output.type = "text"

completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}

merged_data <- readRDS(file="E:/CE Survey/Data/Processed/merged_data_Oct272017.rds") # file created by 10252017_kevin_comments.R

merged_data <- merged_data[merged_data$interest_rate_cur==merged_data$interest_rate_org,]

merged_data <- merged_data[!is.na(merged_data$NEWID.x),]

merged_data['total_exp_less_mtg'] <- merged_data$total_exp - merged_data$mtg_pmt_calc_mor*3



merged_data$income_class <- ifelse(merged_data$income_class==1,4000,
                                   ifelse(merged_data$income_class==2,8000,
                                          ifelse(merged_data$income_class==3,12000,
                                                 ifelse(merged_data$income_class==4,17000,
                                                        merged_data$income_class*10000-25000))))

merged_data <- merged_data[merged_data$total_exp>100 & merged_data$hours_worked_ref>0 & !is.na(merged_data$total_exp),]
merged_data <- merged_data[merged_data$total_exp <= quantile(merged_data$total_exp,0.99,na.rm = TRUE) & merged_data$total_exp >= quantile(merged_data$total_exp,0.01,na.rm = TRUE),]


merged_data['housing_exp_calc'] <- merged_data$major_appliances+merged_data$small_appliances + merged_data$maintainance_repairs+merged_data$house_furnishings_and_equip+merged_data$mtg_principal_outlay

merged_data['non_durable'] <- merged_data$food_exp+ merged_data$entertainment_exp + merged_data$utility_exp + merged_data$transport_outlays - merged_data$vehicle_purchase_outlay + merged_data$apperel_and_services_exp+merged_data$miscellaneous_exp 

merged_data['durable'] <- merged_data$major_appliances + merged_data$small_appliances + merged_data$vehicle_purchase_outlay

merged_data['services'] <- merged_data$education_exp + merged_data$domestic_services + merged_data$health_exp


income_medians <- ddply(merged_data,.(state,interview_year),summarise,q1_income = quantile(income_class,0.25,na.rm=TRUE),median_income = median(income_class,na.rm=TRUE),q3_income = quantile(income_class,0.75,na.rm=TRUE))

merged_data <- merge(merged_data,income_medians,by=c("state","interview_year"),all.x = TRUE)

merged_data$mtg30us <- merged_data$mtg30us*100

```

# Descriptives
```{r descs}

stargazer(merged_data[, c("mpay_mtg","calc_mtg_pmt","non_durable","services","durable","org_year","interview_year","income_class","family_size","age_of_ref","total_exp","total_exp_less_mtg","food_exp","all_major_outlays","entertainment_exp","vehicle_purchase_outlay","major_appliances","small_appliances","house_furnishings_and_equip","maintainance_repairs","utility_exp")], type = output.type, summary.stat = c("mean", "sd", "p25", "median", "p75", "n"),notes = "",digits = 4)
```

# Balancing Tests
```{r balancing}
formula_text <- "~0|org_state+interview_state|(mpay_mtg~mtg30us)|interview_year+CUID"

variables <- c("log(income_class)","age_of_ref","family_size","log(mtg_amount)")
i = 1
felm_regs <- list()
for(variable in variables) {
  # print(i)
  felm_regs[[i]] <- felm(as.formula(paste(variable,formula_text,sep="")),data=merged_data)
  i=i+1
}

stargazer(felm_regs,type = output.type,no.space = TRUE,column.labels = gsub("_"," ",variables))
```

# Kernal Regression
```{r kernal_reg}

library(np)
temp <- merged_data

temp <- completeFun(temp,c("mtg30us","mpay_mtg","org_state","interview_state"))

adj_mpay_mtg <- felm(mpay_mtg~0|factor(org_state)+factor(interview_state),data=temp)
adj_mtg30us <- felm(mtg30us~0|factor(org_state)+factor(interview_state),data=temp)
temp['resid_mpay_mtg'] <- adj_mpay_mtg$residuals
temp['resid_mtg30us'] <- adj_mtg30us$residuals

bw <- npregbw(formula=as.vector(temp$resid_mpay_mtg)~as.vector(temp$resid_mtg30us),bws=(max(temp$resid_mtg30us)-min(temp$resid_mtg30us))/10,bwtype="fixed",bandwidth.compute=F)

model <- npreg(bws=bw,gradients = TRUE)
plot.out <- plot(model, plot.errors.method="asymptotic",plot.errors.style="band",plot.behavior="data")
y.eval <- fitted(plot.out$r1)
x.eval <- plot.out$r1$eval[,1]
y.se <- se(plot.out$r1)
y.lower.ci <- y.eval+1.96*y.se[,1]
y.upper.ci <- y.eval+1.96*y.se[,2]
df <- as.data.frame(cbind(x.eval,y.eval,y.lower.ci,y.upper.ci))

df <- melt(df,id="x.eval")

g2 <- ggplot(df,aes(x=x.eval, y=value,colour=variable)) + geom_line(aes(linetype=variable), size=1) +scale_linetype_manual(values = c(1,2,2))+scale_colour_manual(values=c("black","gray40","gray40"))+ theme_bw()+ylab("FE adjusted mortgage-to-loan") + xlab("FE adjusted IV")+ labs(title = "")+ theme(legend.position="none")+scale_y_continuous(limits = c(-4e-4, 5e-4))
g2



```


# Consumption Regressions - Main Categories

Non-durable: food + entertainment + utility + transport + apperal <br/>
Durable : Major appliances + small appliances + vehicle purchase outlays <br/>
Services: Education + Domestic Services + Health <br/>
housing_exp_calc : major_appliances+small_appliances + maintainance_repairs+house_furnishings_and_equip<br/>
```{r regs_main_cats}
endo_var = "mpay_mtg"
instrument = "mtg30us"
cluster_var = "interview_year+state+CUID"

controls <- "log(income_class)+family_size+I(age_of_ref^2)|org_state+interview_state"



dep_vars <- c("non_durable","durable","services","housing_exp_calc","total_exp")#,"food_exp","all_major_outlays","entertainment_exp","transport_outlays","vehicle_purchase_outlay","major_appliances","small_appliances","house_furnishings_and_equip","maintainance_repairs","utility_exp")


for(dep_var in dep_vars) {
  cat(paste("\n\n\n*******************",dep_var,"*********************\n\n\nSummary Statistics\n"))
  reg_data <- merged_data
  eval(parse(text=paste("reg_data$",dep_var," <- reg_data$",dep_var,"+1",sep="")))
  eval(parse(text=paste("reg_data$",dep_var," <- ifelse(is.na(reg_data$",dep_var,"),1,reg_data$",dep_var,")",sep="")))
  eval(parse(text=paste("reg_data['",dep_var,"_dummy'] <- ifelse(reg_data$",dep_var,">1,1,0)",sep="")))
  
  log_dep_var = paste("log(",dep_var,")",sep="")
  dummy_dep_var = paste(dep_var,"_dummy",sep="")
  
  if(dep_var %in% c("total_exp_less_mtg")) {
    eval(parse(text=paste("reg_data <- reg_data[reg_data$",dep_var,">1,]",sep="")))
  }

  regs <- list()
  regs[[1]] <- felm(as.formula(paste(log_dep_var,"~",endo_var,"+",controls,"|0|",cluster_var,sep="")),data = reg_data)
  regs[[2]] <- felm(as.formula(paste("I(",endo_var,"*1e5)~",instrument,"+",controls,"|0|",cluster_var,sep="")),data = reg_data)
  regs[[3]] <- felm(as.formula(paste(log_dep_var,"~",controls,"|(",endo_var,"~",instrument,")|",cluster_var,sep="")),data = reg_data)
  regs[[4]] <- felm(as.formula(paste(dep_var,"~",controls,"|(",endo_var,"~",instrument,")|",cluster_var,sep="")),data = reg_data)
  regs[[5]] <- felm(as.formula(paste(log_dep_var,"~",controls,"|(",endo_var,"~",instrument,")|",cluster_var,sep="")),data = reg_data[reg_data$income_class < reg_data$median_income,])
  regs[[6]] <- felm(as.formula(paste(log_dep_var,"~",controls,"|(",endo_var,"~",instrument,")|",cluster_var,sep="")),data = reg_data[reg_data$income_class >= reg_data$median_income,])
  
  eval(parse(text=paste("print(summary(reg_data$",dep_var,"-1))",sep="")))

  
  stargazer(regs,type="text",dep.var.labels.include = FALSE,omit.stat = c("f","rsq","ser"),
            column.labels = c("Log","First Stage","Log","Level","Log - Low Income", "Log - High Income"),
            add.lines = list(c("Cond. F","","",round(condfstat(regs[[3]])[[1]],2),round(condfstat(regs[[4]])[[1]],2)
                               ,round(condfstat(regs[[5]])[[1]],2),round(condfstat(regs[[6]])[[1]],2))),title = gsub("_"," ",dep_var))
  
}

```


# Consumption Regressions - Sub Categories
```{r regs_sub_cats}
endo_var = "mpay_mtg"
instrument = "mtg30us"
cluster_var = "interview_year+state+CUID"

controls <- "log(income_class)+family_size+I(age_of_ref^2)|org_state+interview_state"


dep_vars <- c("mtg_principal_outlay","food_away","food_exp","entertainment_exp","utility_exp","transport_outlays","apperel_and_services_exp","major_appliances","small_appliances","vehicle_purchase_outlay","education_exp","domestic_services","health_exp")


for(dep_var in dep_vars) {
  cat(paste("\n\n\n*******************",dep_var,"*********************\n\n\nSummary Statistics\n"))
  reg_data <- merged_data
  eval(parse(text=paste("reg_data$",dep_var," <- reg_data$",dep_var,"+1",sep="")))
  eval(parse(text=paste("reg_data$",dep_var," <- ifelse(is.na(reg_data$",dep_var,"),1,reg_data$",dep_var,")",sep="")))
  eval(parse(text=paste("reg_data['",dep_var,"_dummy'] <- ifelse(reg_data$",dep_var,">1,1,0)",sep="")))
  
  log_dep_var = paste("log(",dep_var,")",sep="")
  dummy_dep_var = paste(dep_var,"_dummy",sep="")
  
  if(dep_var %in% c("total_exp_less_mtg")) {
    eval(parse(text=paste("reg_data <- reg_data[reg_data$",dep_var,">1,]",sep="")))
  }

  regs <- list()
  regs[[1]] <- felm(as.formula(paste(log_dep_var,"~",endo_var,"+",controls,"|0|",cluster_var,sep="")),data = reg_data)
  regs[[2]] <- felm(as.formula(paste("I(",endo_var,"*1e5)~",instrument,"+",controls,"|0|",cluster_var,sep="")),data = reg_data)
  regs[[3]] <- felm(as.formula(paste(log_dep_var,"~",controls,"|(",endo_var,"~",instrument,")|",cluster_var,sep="")),data = reg_data)
  regs[[4]] <- felm(as.formula(paste(dep_var,"~",controls,"|(",endo_var,"~",instrument,")|",cluster_var,sep="")),data = reg_data)
  regs[[5]] <- felm(as.formula(paste(dummy_dep_var,"~",controls,"|(",endo_var,"~",instrument,")|",cluster_var,sep="")),data = reg_data)
  

  eval(parse(text=paste("print(summary(reg_data$",dep_var,"-1))",sep="")))

  
  stargazer(regs,type="text",dep.var.labels.include = FALSE,omit.stat = c("f","rsq","ser"),
            column.labels = c("Log","First Stage","Log","Level","Dummy"),
            add.lines = list(c("Cond. F","","",round(condfstat(regs[[3]])[[1]],2),round(condfstat(regs[[4]])[[1]],2)
                               ,round(condfstat(regs[[5]])[[1]],2))),title = gsub("_"," ",dep_var))
  
}

```

